Popularity Adjusted Block Models are Generalized Random Dot Product Graphs

### 2022-02-03 summary_minh_edited.tex
Premerge

\begin{definition}[Generalized Random Dot Product Graph]
\label{grdpg}
Let $p \geq 1$ and $q \geq 0$ be integers. Define $I_{p,q}$ as the
block diagonal matrix $I_{p,q} = \Bigl[\begin{smallmatrix} I_{p} & 0
  \\ 0 & - I_{q} \end{smallmatrix} \Bigr]$ where $I_{p}$ and $I_{q}$
are the identity matrices of dimensions $p \times p$ and $q \times q$
respectively. Next define $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ where $d = p + q$. 
Let $X = \bigl[x_1 \mid \cdots \mid x_n\bigr]^{\top}$ be
a $n \times d$ matrix with rows $x_i \in \mathcal{X}$. A graph $G$
with adjacency matrix $A$ is said to be a generalized random dot
product graph with latent positions $X$, sparsity parameter $\rho_n
\in (0,1]$
and signature $(p,q)$ if $A \sim \BG(P)$ where the edge probability
matrix $P$ is given
by $P = \rho_n X I_{p,q} X^{\top}$, i.e., the entries of $P$ are of
the form

$$P_{ij} = \rho_n x_i^{\top} I_{p,q} x_j.$$

\end{definition}
We will use the notation \(A \sim \GRDPG_{p,q}(X;\rho_n)\) to denote a random
adjacency matrix \(A\) drawn from latent positions \(X\), sparsity
parameter $\rho_n$ and signature
\((p, q)\).

\begin{remark}
  \label{rem:ase}
We can use Adjacency Spectral Embedding (ASE) 
\citep{doi:10.1080/01621459.2012.699795} to recover the latent vectors of a GRDPG. 
This procedure consists of taking the spectral decomposition of $A$ 
(or $P$ if available) and choosing the $p$ most positive and $q$ most negative
eigenvalues, $D = \diag(\lambda_1, ..., \lambda_p, \lambda_{n - q + 1}, ..., \lambda_n)$, and their corresponding eigenvectors, $V \in \mathbb{R}^{n \times (p + q)}$,
to construct the embedding $\hat{Z} = V |D|^{1/2}$. 
\end{remark}

\begin{definition}[Indefinite Orthogonal Group]
The indefinite orthogonal group with signature $(p, q)$ is
the set $\{Q \in \mathbb{R}^{d \times d} \colon Q I_{p, q} Q^{\top} = I_{p, q}\}$,
denoted as $\mathbb{O}(p, q)$. Here $d = p + q$. 
\end{definition}

@@

>>>>

\begin{definition}[Generalized Random Dot Product Graph]
\label{grdpg}
Let $p \geq 1$ and $q \geq 0$ be integers. Define $I_{p,q}$ as the
block diagonal matrix $I_{p,q} = \Bigl[\begin{smallmatrix} I_{p} & 0
  \\ 0 & - I_{q} \end{smallmatrix} \Bigr]$ where $I_{p}$ and $I_{q}$
are the identity matrices of dimensions $p \times p$ and $q \times q$
respectively. Denote $d = p + q$ and let $\mathcal{X}$ be a subset of $\mathbb{R}^d$ such
that, for any $x \in \mathcal{X}$ and $y \in \mathcal{X}$, we have
$x^{\top} I_{p,q} y \in [0,1]$. 
% Next define $\mathcal{X} = \{x, y \in \mathbb{R}^d : x^\top I_{p,q} y \in [0, 1]\}$ where $d = p + q$. 
Let $X = \bigl[x_1 \mid \cdots \mid x_n\bigr]^{\top}$ be
a $n \times d$ matrix with rows $x_i \in \mathcal{X}$. A graph $G$
with adjacency matrix $A$ is said to be a generalized random dot
product graph with latent positions $X$, sparsity parameter $\rho_n
\in (0,1]$
and signature $(p,q)$ if $A \sim \BG(P)$ where the edge probability
matrix $P$ is given
by $P = \rho_n X I_{p,q} X^{\top}$, i.e., the entries of $P$ are of
the form $P_{ij} = \rho_n x_i^{\top} I_{p,q} x_j.$
\end{definition}
We will use the notation \(A \sim \GRDPG_{p,q}(X;\rho_n)\) to denote a random
adjacency matrix \(A\) drawn from latent positions \(X\), sparsity
parameter $\rho_n$ and signature
\((p, q)\).

\begin{definition}[Indefinite Orthogonal Group]
  \label{def:indefinite}
The indefinite orthogonal group with signature $(p, q)$ is
the set $\{Q \in \mathbb{R}^{d \times d} \colon Q I_{p, q} Q^{\top} = I_{p, q}\}$,
denoted as $\mathbb{O}(p, q)$. Here $d = p + q$. 
\end{definition}

@@

\begin{remark}
  \label{rem:ase}
We can use Adjacency Spectral Embedding (ASE) 
\citep{doi:10.1080/01621459.2012.699795} to recover the latent vectors of a GRDPG. 
This procedure consists of taking the spectral decomposition of $A$
and keeping the $p + q$ largest eigenvalues (in modulus) and
corresponding eigenvectors of $A$. More specifically, let $A$ be an $n
\times n$ adjacency matrix and denote the
eigendecomposition of $A$ as
$$A = \sum_{i} \hat{\lambda}_i \hat{v}_i \hat{v}_i^{\top}, \quad
\text{where} \,\,
|\hat{\lambda}|_1 \geq |\hat{\lambda}_2| \geq \dots \geq
|\hat{\lambda}_n|.$$
Now let $\hat{D}$ be a diagonal matrix whose diagonal entries are
the eigenvalues
$\{\hat{\lambda}_i\}_{i=1}^{d}$ arranged in decreasing values (not in
decreasing modulus). Let $\hat{V}$ be the $n \times d$ matrix whose
columns are the corresponding eigenvectors $\{\hat{v}_i\}_{i=1}^{d}$
arranged in the same order as the diagonal entries of $\hat{D}$. Now
define $\hat{Z} = \hat{V} |\hat{D}|^{1/2}$ where the $|\cdot|$
operation is applied elementwise to the entries of $\hat{D}$. Then
$\hat{Z}$ serves as an estimate of $X$, up to some non-identifiable
orthogonal transformation $Q$ as described in
Definition~\ref{def:indefinite} and
Remark~\ref{rem:non_identifiable}. 
% choosing the $p$ most positive and $q$ most negative
% eigenvalues, $D = \diag(\lambda_1, ..., \lambda_p, \lambda_{n - q +
%   1}, ..., \lambda_n)$, and their corresponding eigenvectors, $\hat{V}
% \in \mathbb{R}^{n \times (p + q)}$, to construct the embedding
%$\hat{Z} = V |D|^{1/2}$. More specifically, 
\end{remark}
